---
title: "BIS 15L Midterm 2"
author: Ricardo Pineda
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code should be organized, clean, and run free from errors. Be sure to **add your name** to the author header above. You may use any resources to answer these questions (including each other), but you may not post questions to Open Stacks or external help sites. There are 10 total questions.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! Your plots should use consistent aesthetics throughout.  

This exam is due by **12:00p on Tuesday, February 22**.  

## Gapminder
For this assignment, we are going to use data from  [gapminder](https://www.gapminder.org/). Gapminder includes information about economics, population, social issues, and life expectancy from countries all over the world. We will use three data sets, so please load all three as separate objects.    

1. population_total.csv  
2. income_per_person_gdppercapita_ppp_inflation_adjusted.csv  
3. life_expectancy_years.csv  

```{r}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
library(ggthemes)
```

```{r}
pop_total <- readr:: read_csv("data/population_total.csv")
income_and_inflation <- readr:: read_csv("data/income_per_person_gdppercapita_ppp_inflation_adjusted.csv")
life_exp <- readr:: read_csv("data/life_expectancy_years.csv ")
```
```{r}
glimpse(pop_total)
tibble(pop_total)
naniar::miss_var_summary(pop_total)
```
```{r}
glimpse(income_and_inflation)
tibble(income_and_inflation)
naniar::miss_var_summary(income_and_inflation)
```
```{r}
glimpse(life_exp)
tibble(life_exp)
naniar::miss_var_summary(life_exp)
```



1. (3 points) Once you have an idea of the structure of the data, please make each data set tidy (hint: think back to pivots) and store them as new objects. You will need both the original (wide) and tidy (long) data!  

```{r}
tidy_pop_total <- pop_total %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "population_total")
tidy_pop_total
```
```{r}
tidy_income_and_inflation <- income_and_inflation %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "income")
tidy_income_and_inflation
```
```{r}
tidy_life_exp <- life_exp %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "life_expectancy") %>%
    filter(!is.na(life_expectancy))
tidy_life_exp
```


2. (1 point) How many different countries are represented in the data? Provide the total number and their names. Since each data set includes different numbers of countries, you will need to do this for each one.  

I. For pop total

```{r}
tidy_pop_total %>%
  summarize(n_countries = n_distinct(country))
#for total number of countries
```
```{r}
pop_total %>%
  select(country) %>%
  group_by(country)
#list of names of the different countries
```

II. For income and inflation

```{r}
tidy_income_and_inflation %>%
  summarize(n_countries = n_distinct(country))
#for total number of countries
```
```{r}
income_and_inflation %>%
  select(country) %>%
  group_by(country)
#list of names of the different countries
```

III. For life exp

```{r}
tidy_life_exp %>%
  summarize(n_countries = n_distinct(country))
#for total number of countries
```
```{r}
life_exp %>%
  select(country) %>%
  group_by(country)
#list of names of the different countries
```

## Life Expectancy  

3. (2 points) Let's limit the data to 100 years (1920-2020). For these years, which country has the highest average life expectancy? How about the lowest average life expectancy?  

```{r}
tidy_life_exp %>%
  names()
```

```{r}
tidy_life_exp %>%
  filter(between(year, 1920,2020)) %>%
  group_by(country) %>%
  summarize(mean_life_expectancy = mean(life_expectancy)) %>%
  arrange(desc(mean_life_expectancy)) %>%
  head(10)

#For the top 10 highest average life expectancy
```
```{r}
tidy_life_exp %>%
  filter(between(year, 1920,2020)) %>%
  group_by(country) %>%
  summarize(mean_life_expectancy = mean(life_expectancy)) %>%
  arrange((mean_life_expectancy)) %>%
  head(10)

#For the top 10 lowest average life expectancy
```
4. (3 points) Although we can see which country has the highest life expectancy for the past 100 years, we don't know which countries have changed the most. What are the top 5 countries that have experienced the biggest improvement in life expectancy between 1920-2020? 

```{r}
#Comparing the min and max of the time interval
tidy_life_exp %>%
  filter(between(year, 1920,2020)) %>%
  group_by(country) %>%
  summarize(min_for_interval = min(life_expectancy),
            max_for_interval = max(life_expectancy),
            largest_improvement = max_for_interval - min_for_interval) %>%
  arrange(desc(largest_improvement)) %>%
  head(10)
```
```{r}
#When comparing only the years 1920 and 2020 

tidy_life_exp %>%
  filter(year == 1920 | year == 2020) %>%
  group_by(country) %>%
  summarize(min_for_interval = min(life_expectancy),
            max_for_interval = max(life_expectancy),
            largest_change = max_for_interval - min_for_interval) %>%
  arrange(desc(largest_change)) %>%
  head(10)

```

5. (3 points) Make a plot that shows the change over the past 100 years for the country with the biggest improvement in life expectancy. Be sure to add appropriate aesthetics to make the plot clean and clear. Once you have made the plot, do a little internet searching and see if you can discover what historical event may have contributed to this remarkable change.

```{r}
tidy_life_exp %>%
  filter(country == "Kuwait") %>%
  filter(between(year, 1922,2022)) %>%
  ggplot(aes(x=year, y= life_expectancy)) +
  geom_point(shape=1)+
  scale_x_discrete(breaks=seq(1920, 2025, 5)) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1, vjust = 1))+
  labs(title = "Life Expectancy in Kuwait Over Time",
       x = "Year",
       y = "Life Expectancy")

#"Between 1946 and 1982, Kuwait experienced a period of prosperity driven by oil and its liberal atmosphere.In popular discourse, the years between 1946 and 1982 are referred to as the "Golden Era of Kuwait". - Wikipedia

```


## Population Growth
6. (3 points) Which 5 countries have had the highest population growth over the past 100 years (1920-2020)?  

```{r}
#Comparing the min and max of the time interval
tidy_pop_total %>%
  filter(between(year, 1920, 2020)) %>%
  group_by(country) %>%
  summarise(min_for_interval = min(population_total),
            max_for_interval = max(population_total),
            largest_improvement = max_for_interval - min_for_interval) %>%
  arrange(desc(largest_improvement)) %>%
  head(5)
```
```{r}
#When comparing only the years 1920 and 2020 
tidy_pop_total %>%
  filter(year == 1920 | year == 2020) %>%
  group_by(country) %>%
  summarise(min_for_interval = min(population_total),
            max_for_interval = max(population_total),
            largest_improvement = max_for_interval - min_for_interval) %>%
  arrange(desc(largest_improvement)) %>%
  head(5)
```



7. (4 points) Produce a plot that shows the 5 countries that have had the highest population growth over the past 100 years (1920-2020). Which countries appear to have had exponential growth?  

```{r}
tidy_pop_total %>%
  filter(country == "India" |country == "China" | country == "Indonesia" | country == "United States" | country == "Pakistan" ) %>%
  filter(between(year, 1922,2022)) %>%
  ggplot(aes(x=year, y= population_total, color =  country)) +
  geom_point(shape=1)+
  scale_x_discrete(breaks=seq(1920, 2025, 5)) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1, vjust = 1))+
  labs(title = "Population Over Time",
       x = "Year",
       y = "Population Total")

#China and India with exponential growth
```


## Income
The units used for income are gross domestic product per person adjusted for differences in purchasing power in international dollars.

8. (4 points) As in the previous questions, which countries have experienced the biggest growth in per person GDP. Show this as a table and then plot the changes for the top 5 countries. With a bit of research, you should be able to explain the dramatic downturns of the wealthiest economies that occurred during the 1980's.

```{r}
tidy_income_and_inflation %>%
  filter(between(year, 1922,2022)) %>%
  group_by(country) %>%
  summarise(min_for_interval = min(income),
            max_for_interval = max(income),
            largest_improvement = max_for_interval - min_for_interval) %>%
  arrange(desc(largest_improvement)) %>%
  head(5)
```

```{r}
#Cut the time frame to 1950 to 2050  to make plot easier to read
tidy_income_and_inflation %>%
  filter(between(year, 1950,2050)) %>%
  filter(country == "United Arab Emirates" | country == "Qatar" | country == "Brunei" | country == "Kuwait" | country == "Luxembourg") %>%
  group_by(year, country) %>%
  ggplot(aes(x=year, y= income, group = country, color =  country)) +
  geom_line() +
  geom_point(shape=1)+
  scale_x_discrete(breaks=seq(1900, 2020, 10)) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1, vjust = 1))+
  labs(title = "Income Over Time",
       x = "Year",
       y = "Income")
```





9. (3 points) Create three new objects that restrict each data set (life expectancy, population, income) to the years 1920-2020. Hint: I suggest doing this with the long form of your data. Once this is done, merge all three data sets using the code I provide below. You may need to adjust the code depending on how you have named your objects. I called mine `life_expectancy_100`, `population_100`, and `income_100`. For some of you, learning these `joins` will be important for your project.  

```{r}
population_100 <- pop_total %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "population_total")%>%
  filter(between(year, 1920,2020))
population_100
```

```{r}
income_100 <- income_and_inflation %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "income") %>%
  filter(between(year, 1920,2020))
income_100
```
```{r}
life_expectancy_100 <- life_exp %>%
  pivot_longer(-c(country),
              names_to = "year",
              values_to = "life_expectancy") %>%
  filter(between(year, 1920,2020))
life_expectancy_100
```





```{r}
gapminder_join <- inner_join(life_expectancy_100, population_100, by= c("country", "year"))
gapminder_join <- inner_join(gapminder_join, income_100, by= c("country", "year"))
gapminder_join
```

10. (4 points) Use the joined data to perform an analysis of your choice. The analysis should include a comparison between two or more of the variables `life_expectancy`, `population`, or `income.`  

```{r}
#Population vs income over time
#Plot illustrates how over time life expectancy throughout the world steadily increases. Wealthier countries also have a higher life expectancy but income difference make less of an impact in more recent decades
pop_vs_inc <- gapminder_join %>%
  filter(year == 1900 | year == 1920 | year == 1940 | year == 1960 | year == 1980 | year == 2000 | year == 2020) %>%
  filter(!is.na(life_expectancy)) %>%
   ggplot(aes(x = life_expectancy,
              y= income,
              color = income))+
      geom_point(na.rm=T) +
      scale_y_log10() +
      scale_x_log10() +
      facet_wrap(~year)+
      geom_smooth(method=lm, se=T, color = "red", alpha = 0.1)+
      theme(axis.text.x = element_text(angle = 40, hjust = 1, vjust = 1))+
      labs(title = "Income vs Life Expectancy Over Time",
       x = "Life Expectancy (yrs)",
       y= "Income")

pop_vs_inc
  

```

