---
title: "Lab 14 HW"
output: 
  html_document: 
    keep_md: yes
author: "your name here"
date: '2022-02-28'
---


### Load the Libraries
```{r}
library(tidyverse)
library(janitor)
library("palmerpenguins")

#install.packages("ggVennDiagram")
library(ggVennDiagram)
library(RColorBrewer)

#install.packages("ggworldcloud")
library(ggwordcloud)


options(scipen=999) #cancels the use of scientific notation for the session
```


### Data
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))

beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")
```

### 1. 
Clean up the column names (no capitals, not spaces) of `superhero_info`, then use 2 functions to remind yourself of structure of the `superhero_info` data set.

```{r}
clean_superhero_info <- clean_names(superhero_info)
tibble(clean_superhero_info)
glimpse(clean_superhero_info)
```

### 2.
Are bad guys bigger? Make box-plots of weight by `alignment` (alignment on the x-axis).
```{r}
clean_superhero_info %>% 
  filter(!is.na(alignment)) %>%
  ggplot(aes(x = alignment, y = weight, fill = alignment)) +
  geom_boxplot(na.rm = TRUE)
```

### 3. 
Now, make a violin plot of weight by `alignment` (alignment on the x-axis). Add some color!
  What information can you observe in the violin plot that was not visible in the boxplot?
```{r}
clean_superhero_info %>% 
  filter(!is.na(alignment)) %>%
  ggplot(aes(x = alignment, y = weight, fill = alignment)) +
 geom_violin(alpha = .5, na.rm = TRUE)

#You can see what the distributions look like
#Ex: two peaks for good guys
```



### 4. 
Use `alpha = .5` in `geom_boxplot()` and `geom_violin()` to layer both plots on top of one another. What does this tell you about the distribution of weight in "`bad`" guys?
```{r}
clean_superhero_info %>% 
  filter(!is.na(alignment)) %>%
  ggplot(aes(x = alignment, y = weight, fill = alignment)) +
  geom_boxplot(na.rm = TRUE, alpha = .5) +
  geom_violin(alpha = .5, na.rm = TRUE)

#There is a long tail for the distribution of weight of bad guys and there are more bad guys in the tail than there are good guys in the good guys distribution.
```

### 5. 
Box plots are great for showing how the distribution of a numeric variable (e.g. weight) varies among a categorical variable (e.g. alignment).
  Make your own violin plot with the superhero data. 
  What is your numeric data? 
  What is your categorical variable?

```{r}
clean_superhero_info %>% 
  ggplot(aes(x = gender, y = height, fill = gender)) +
  geom_boxplot(na.rm = TRUE, alpha = .5) +
  geom_violin(alpha = .5, na.rm = TRUE)

#Numerical data is height
#Categorical data is gender
```

### 6. 
Remind yourself what `beachbugs` looks like. Then generate a heatmap of buglevels by site and year. 
color it with `scale_fill_gradient(low="yellow", high="red")` or colors of your choice. you may find it looks best with one color!
(dont forget, `coord_flip()` is a quick way to improve the look of your plot if you dont like the default orientation)

```{r}
glimpse(beachbugs_long)
tibble(beachbugs_long)
```

```{r}
beachbugs_long %>%
  ggplot(aes( x = year, y = site, fill= buglevels)) +
  geom_tile() + 
  scale_fill_gradient(low="yellow", high="red")
```

### 7.  
Use the provided code to normalize the beachbug data set. 
Then make a heatmap with the `beachbugs_normalized` data, and use the same color chois as above. Which heatmap do you think is more informative? why?

```{r}
#makes a new column of the highest buglevel observed at each site
beachbugs_w_max <- beachbugs_long %>% 
  group_by(site) %>% 
  mutate(max_buglevel = max(buglevels, na.rm=T)) %>% 
  arrange(site)
beachbugs_w_max

#makes a new table where the buglevels are normalized to the max_buglevel
beachbugs_normalized <- beachbugs_w_max %>% 
  group_by(site) %>% 
  mutate(norm_buglevel = buglevels/max_buglevel) %>% 
  arrange(site,year) %>%
  select(site, year, norm_buglevel)   # you dont have to select(), but I think its a clearer looking table
beachbugs_normalized
```
 
```{r}
beachbugs_w_max %>%
  ggplot(aes( x = year, y = site, fill= buglevels)) +
  geom_tile() + 
  scale_fill_gradient2(low="green", mid = "white", high="red")
```
```{r}
beachbugs_normalized %>%
  ggplot(aes( x = year, y = site, fill= norm_buglevel)) +
  geom_tile() + 
  scale_fill_gradient2(low="green", mid = "white", high="red")
```

I think the normalized heatmap is more informative as it shows what years and where there is a higher amount of bugs than normal while the other heatmap just shows the highest.

### 8.
Let's make a venn diagram of `superhero_info`, from 4 questions:
Is their alignment evil? 
Are their eyes red?
Are they male?
Are they bald?

Start by making the 4 vectors, then the list of vectors. The vector for alignment is provided:
### super heros venn
```{r}
# evil

evil_vec <- clean_superhero_info %>%
  filter(alignment == "bad")%>%
  pull(name)

# red eyes
red_eyes_vec <- clean_superhero_info %>%
  filter(eye_color == "red")%>%
  pull(name)

# male
male_vec <- clean_superhero_info %>%
  filter(gender == "Male")%>%
  pull(name)

# bald
bald_vec <- clean_superhero_info %>%
  filter(hair_color == "No Hair")%>%
  pull(name)




```

Your list of vectors will look something like this:
```{r}

evil_red_male_bald_list <- list(evil_vec, red_eyes_vec, male_vec, bald_vec)


```

### 9. 
Let's make the venn diagram! use the code from lab as a reference. 
```{r}
# something like:
ggVennDiagram( evil_red_male_bald_list, category.names = c("Evil", "Red Eyes", "Male", "Bald"))
```


### 10. Choose one intersection of the venn diagram that is interesting to you. Use dplyr to find the names of the superheros in that intersection. 

```{r}
#DC comics
dc_vec <- clean_superhero_info %>%
  filter(publisher == "DC Comics")%>%
  pull(name)

# female
female_vec <- clean_superhero_info %>%
  filter(gender == "Female")%>%
  pull(name)

# Black hair
black_hair_vec <- clean_superhero_info %>%
  filter(hair_color == "Black")%>%
  pull(name)


dc_female_black_hair_list <- list(dc_vec, female_vec, black_hair_vec)
```



### 11. Make another venn diagram with the `superhero_info` data. What are your questions? ( At least 2!) 
```{r}
ggVennDiagram(dc_female_black_hair_list, category.names = c("DC","Female", "Black Hair"))

#How many of the DC characters are female?
#60
#How many superheros are female and have black hair?
#46
```



### 12.
What are some very common super powers? Lets make a word cloud with the `superhero_powers` data.

Use the provided code to make the frequency table, then make a word cloud with it. 
Remember, you can change `scale_size_area(max_size = 20)` to a different number if the words wont fit!

```{r}
head(superhero_powers)
```
```{r}
power_frequency <- superhero_powers %>% 
  select(-hero_names) %>%
  summarise_all(~(sum(.))) %>%
  pivot_longer(everything(), names_to = "power", values_to = "freq")
power_frequency
```


```{r}
power_frequency %>% 
ggplot(aes(
  label = power, 
  size = freq,
  color = power
  )) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 10)


```
### 13.  
Who are some very powerful supers? 
 Lets make a different word cloud with the `superhero_powers` data. 
Use the provided code to make the frequency table, then make a word cloud with it.
You can use `hero_names` for the labels, and `sum_powers` for size. 

```{r}

power_quantity <- superhero_powers %>% 
  pivot_longer(-hero_names, names_to = "power", values_to = "yes_or_no") %>% 
  group_by(hero_names) %>% 
  mutate(sum_powers = sum(yes_or_no, na.rm=T)) %>%
  arrange(desc(sum_powers), hero_names, desc(yes_or_no))
power_quantity

#we have to trim down to only the top 50, or it will crowd the image!
power_quantity <- power_quantity %>% 
  ungroup %>%
  distinct(hero_names, sum_powers) %>%
  slice_max(sum_powers, n = 50)
power_quantity

```



```{r}
power_quantity %>% 
ggplot(aes(
  label = hero_names, 
  size = sum_powers,
  color = hero_names)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 10)

```

## That's it! 🎉
Thanks for coding with us all winter! 
Now go finish your group project :) 

